---
title: "Bikeshare Usage and Equity in Washington DC"
author: "Yuetong Wang"
date: "`r Sys.Date()`"
output:
   html_document:
     toc: TRUE
     toc_float: TRUE
     theme: "spacelab"
     code_folding: show
     number_section: TRUE
     fig_height: 4
     fig_width: 6
     fig_align: 'center'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(readr)
library(tidyverse)
library(corrplot)
library(car)
library(PerformanceAnalytics)
library(leaps)
library(sf)
library(gstat)
library(rpart)
library(rpart.plot)
library(ranger)
library(nlme)
```

# Load the dataset

```{r}
Final_Table <- read_csv("./Final Table.csv", show_col_types = FALSE)
Final_Table
```

# Correlation between variables

```{r fig.width=4, fig.height=3}
corr_plot <- cor(Final_Table[,6:20])
corrplot(corr_plot, method = "circle", tl.col="black", tl.cex=0.8)
```

```{r fig.width=4, fig.height=3}
corrplot(corr_plot, method = "number", number.cex = 0.7, tl.col="black", tl.cex=0.8)
```

```{r fig.width=4, fig.height=3}
chart.Correlation(Final_Table[,6:20], histogram=TRUE, pch=19)
```

# Make a Spatial Dataframe

```{r}
Final_Table_sf <- st_as_sf(Final_Table, coords = c("lng","lat"))
class(Final_Table_sf)
st_crs(Final_Table_sf) <- 4326
```

```{r}
ggplot(data = Final_Table_sf) +
  geom_sf(mapping = aes(size=MHI, color=O_Trip)) +
  labs(x = "Longitude", y = "Latitude", title = "Trip per year", color="Bikeshare Trips", size = "MHI")
```

# Multiple Linear Regression

## OLS Model including all variables

```{r}
# Select the dependent and independent variables
Final_Table_model <- Final_Table[,6:20]
```

```{r}
Linear_model <- lm(L_O_Trip ~ SWalk_Den + BikeL_Den + Capacity + N_OBike + N_MetroS + N_BusS + Pop_Den + P_White + P_Young + MHI + P_NoCar + P_Renter + Unemploy_R + Crash, data = Final_Table_model)
summary(Linear_model)
plot(Linear_model)
```


```{r}
# Calculate the VIF values
vif(Linear_model)

# Calculate the AIC value
AIC(Linear_model)

```

## Best Subset Selection including all variables

```{r}
regfit_full = regsubsets(L_O_Trip ~ SWalk_Den + BikeL_Den + Capacity + N_OBike + N_MetroS + N_BusS + Pop_Den + P_White + P_Young + MHI + P_NoCar + P_Renter + Unemploy_R + Crash, data = Final_Table_model, nvmax = 14)
summary(regfit_full)
```

```{r}
reg_summary = summary(regfit_full)
reg_summary$adjr2
reg_summary$bic
```


## OLS not include P_White

```{r}
Linear_model_NoPWhite <- lm(L_O_Trip ~ SWalk_Den + BikeL_Den + Capacity + N_OBike + N_MetroS + N_BusS + Pop_Den + P_Young + MHI + P_NoCar + P_Renter + Unemploy_R + Crash, data = Final_Table_model)
summary(Linear_model_NoPWhite)
plot(Linear_model_NoPWhite)
```

```{r}
# Calculate the VIF values
vif(Linear_model_NoPWhite)

# Calculate the AIC value
AIC(Linear_model_NoPWhite)

BIC(Linear_model_NoPWhite)

```

### Best Subset Selection not include P_White

```{r}
regfit_full = regsubsets(L_O_Trip ~ SWalk_Den + BikeL_Den + Capacity + N_OBike + N_MetroS + N_BusS + Pop_Den + P_Young + MHI + P_NoCar + P_Renter + Unemploy_R + Crash, data = Final_Table_model, nvmax = 13)
summary(regfit_full)
```

### Check the adjusted R-square

```{r}
reg_summary = summary(regfit_full)
reg_summary$adjr2
reg_summary$bic
```

# Generalized Least Squares

```{r}
Final_Table_sf$residual <- Linear_model_NoPWhite$residuals

v <- variogram(residual~1,data=Final_Table_sf)
plot(v)
```


```{r}
# change the crs
Final_Table_sf_utm <- st_transform(Final_Table_sf, crs=26918)
st_bbox(Final_Table_sf_utm)

Final_Table_utm <- cbind(Final_Table_sf_utm, st_coordinates(Final_Table_sf_utm))
Final_Table_utm
names(Final_Table_utm)

model_gls <- gls(
  model = L_O_Trip ~ SWalk_Den + BikeL_Den + Capacity + N_OBike + N_MetroS + N_BusS + Pop_Den + P_Young + MHI + P_NoCar + P_Renter + Unemploy_R + Crash, 
  data = Final_Table_utm,
  method = "ML",
  correlation = corExp(form = ~X + Y,
                       nugget = TRUE,
                       value = c(7000/3,0.4))
)
summary(model_gls)
```

```{r}
plot(Final_Table_utm$L_O_Trip ~ predict(model_gls), pch=20, asp=1, 
     xlab="Fitted by GLS", ylab="Actual", main="log trips")
grid()
abline(0,1)
```


# Building Regression Tree

```{r}
Final_Table_model_tree <- rpart(L_O_Trip ~ SWalk_Den + BikeL_Den + Capacity + N_OBike + N_MetroS + N_BusS + Pop_Den + P_White + P_Young + MHI + P_NoCar + P_Renter + Unemploy_R + Crash, data = Final_Table_model)
print(Final_Table_model_tree)
```
```{r}
rpart.plot(Final_Table_model_tree)
```

```{r}
data.frame(variableImportance = 100 * Final_Table_model_tree$variable.importance / sum(Final_Table_model_tree$variable.importance))
```

## Not include P_White

```{r}
Final_Table_model_tree_NoPWhite <- rpart(L_O_Trip ~ SWalk_Den + BikeL_Den + Capacity + N_OBike + N_MetroS + N_BusS + Pop_Den + P_Young + MHI + P_NoCar + P_Renter + Unemploy_R + Crash, data = Final_Table_model)
print(Final_Table_model_tree_NoPWhite)
```

```{r}
rpart.plot(Final_Table_model_tree_NoPWhite)
```

```{r}
data.frame(variableImportance = 100 * Final_Table_model_tree_NoPWhite$variable.importance / sum(Final_Table_model_tree_NoPWhite$variable.importance))
```


# Random Forest

```{r}
library(randomForest)
library(caret)

# Fit a random forest model
rf_model <- randomForest(L_O_Trip ~ ., data = Final_Table_model)

summary(rf_model)

rf_model

# Get the feature importance scores
importance(rf_model)

# Plot the variable importance
varImpPlot(rf_model)

```

## Random Forest not include P_White

```{r}
# Fit a random forest model
rf_model_NoPWhite <- randomForest(L_O_Trip ~ SWalk_Den + BikeL_Den + Capacity + N_OBike + N_MetroS + N_BusS + Pop_Den + P_Young + MHI + P_NoCar + P_Renter + Unemploy_R + Crash, data = Final_Table_model)

rf_model_NoPWhite

# Get the feature importance scores
importance(rf_model_NoPWhite)

# Plot the variable importance
varImpPlot(rf_model_NoPWhite)
```


